Regression Cube Analysis of Cohort Study Data

### 2015-03-19 draft.tex
wrote Encoding subsec on Regression Cube section

\subsection{Abstracting Regression Results using R$^2$}
\com{R$^2$ are used to encode regression results to color spaces.}
\com{R$^2$ is comparable between regression results of the same type, but not in between types. Therefore it is encoded using different colours.}

@@

\begin{figure}[htb]
 \centering
 \includegraphics[width=2.8in]{figures/placeholder}
 \caption{
 \com{Screenshot of the cube visualization embedded in the framework}
 }
  \label{fig:comparison}
\end{figure}
%%
\com{Simplicity of interface to allow for steep learning curve.}
\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
\com{Cube acts as visual mini map to give a global impression over the data.}
\com{current pane shown using a heat map visualization.}
\com{details on demand for selected regression formula.}
\com{comparison view using reference cubes.}

>>>>

\subsection{Abstracting Regression Results using R$^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.0in]{figures/cube_sketch}
 \caption{
 (a) Overview visualization using a heatmap for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
 %%
 The R$^2$ values extracted from the regression formulas depict the goodness of fit and are mapped to color saturation.
 %%
 A saturated color shows a strong correlation.
 %%
 (b) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ heatmap visualizations.
 %%
 These represent the slices in our cube visualization.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of a overview visualization is to provide a comprehensive view on the data, which is easy to understand.
%% TODO: Cite another paper here!
As described in \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the goodness of fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the R$^2$ value is the metric allowing for this kind of assessment.
%%
Since it is scaled between [0,1] it also allows for comparison between regression models.
%%
We can apply the same information mapping by translating the R$^2$ values to color saturation (Fig.~\ref{fig:Cube} (a)).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g. $Age \sim X + Y$).
%%
Introducing $Z$ creates a 3D \emph{Regression Cube}.
%%
The visualization of R$^2$ values derived from different regression cubes (e.g. $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, the R$^2$ results of different regression methods are encoded using different colors (e.g. blue for linear regression and red for logistic regression).
%%
This way, the cube can easily be extended using other regression types.
%%
For cubes having one fixed target feature, such as $Cancer \sim X + Y + Z$ no such encodings is required and the $z$ dimension can be compared directly.

The goal is to create a overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
Different analysis approaches require different starting points using the \emph{Regression Cube}.

@@

\begin{figure}[htb]
 \centering
 \includegraphics[width=2.8in]{figures/placeholder}
 \caption{
 \com{Screenshots of Cubes from the Evaluation}
 }
  \label{fig:comparison}
\end{figure}
%%
\com{Simplicity of interface to allow for steep learning curve.}
\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
\com{Cube acts as visual mini map to give a global impression over the data.}
\com{current pane shown using a heat map visualization.}
\com{details on demand for selected regression formula.}
\com{comparison view using reference cubes.}
